---
title: "Walmart trip type classification exploratory analysis"
author: "Alex Ding"
date: "December 13, 2015"
output: html_document
---

#Alex's Script on
#Warlmart trip type classification
#Date: 2015 12 13

Read in the data:
```{r}
rm(list=ls())
gc()
library(data.table)
library(zoo)
library(reshape)
setwd("C:/Users/Alex/Desktop/cs/Kaggle/12 27 2015 Walmart tripe type classification/data/")
set.seed(616)
train <- read.csv("train.csv" )
test  <- read.csv("test.csv")
#add an indicator to show whether being purchased or returned
train$Purchase_Ind <- "Purchase"
train[which(train$ScanCount<0),'Purchase_Ind'] <- "Return"
test$Purchase_Ind <- "Purchase"
test[which(test$ScanCount<0),'Purchase_Ind'] <- "Return"
dt.train <- data.table(train)
dt.test <- data.table(test)

```

let's have a first look at the data

```{r}
str(train)
unique(train$TripType)
str(test)
```

NA per column

```{r}
sapply(train,function(x) sum(is.na(x)))
apply(is.na(test),2,sum)
```

Unique values per column
38 types of trip type
69 department in train, 68 department in test
5196 unique FinelineNumber in train
5204 unique FinelineNumber in test
```{r}
sapply(train,function(x) length(unique(x)))
sapply(test,function(x) length(unique(x)))
```


all department description are in the training set
only 5046 FineLineNumber exist in the test set
```{r}
sum(unique(test$DepartmentDescription) %in% unique(train$DepartmentDescription))
sum(unique(test$FinelineNumber) %in% unique(train$FinelineNumber))
```


train has 97714 unique products.
test has 98147 unique products.
there are 71168 Upc codes that are both in train and test

note: when training the model, maybe it is a good idea to transform all 
UPC codes that are 0 or not in the train set to 0, since it has never been seen before

```{r}
length(unique(train[which(!is.na(train$Upc)),]$Upc))
length(unique(test[which(!is.na(test$Upc)),]$Upc))

sum(unique(test[which(!is.na(test$Upc)),]$Upc) %in% unique(train[which(!is.na(train$Upc)),]$Upc))
```

upc definition

first 6 digits are manufacturer code, next 5 digits are item number
last digit is a check digit

department product count by its purchase, return and return/Purchase ratio
The top 5 return/purchase ratio is:
```{r}
xnam = c("DepartmentDescription","Purchase_Ind")
a = data.frame(table(train[xnam]))
a = cast(a,DepartmentDescription~Purchase_Ind)
a = a[with(a, order(-Purchase)), ]

Calculate_Purchase_Return_ratio <- function(in_data){
  in_data['Return'] <- abs(in_data['Return'])
  in_data['RPratioBps']<- in_data['Return']/in_data['Purchase']* 1000
  return (in_data)
}

a = Calculate_Purchase_Return_ratio(a)
head(a,15)
```

department product count by its purchase and return in test
```{r}
xnam = c("DepartmentDescription","Purchase_Ind")
b = data.frame(table(test[xnam]))
b = cast(b,DepartmentDescription~Purchase_Ind)
b = b[with(b, order(-Purchase)), ]
b = Calculate_Purchase_Return_ratio(b)
head(b,15)
```



Let's do the same summary by the Quantities(ScanCOunt) sold 
```{r}
xnam = c("ScanCount","DepartmentDescription","Purchase_Ind")
d<-aggregate(ScanCount ~ ., data=train[, xnam], FUN=sum,order('ScantCount'))
d = cast(d,DepartmentDescription~Purchase_Ind)
d = d[with(d, order(-Purchase)), ]
d[is.na(d)] <- 0
d = Calculate_Purchase_Return_ratio(d)
head(d,15)

xnam = c("ScanCount","DepartmentDescription","Purchase_Ind")
e<-aggregate(ScanCount ~ ., data=test[, xnam], FUN=sum,order('ScantCount'))
e = cast(e,DepartmentDescription~Purchase_Ind)
e = e[with(e, order(-Purchase)), ]
e[is.na(e)] <-0
e = Calculate_Purchase_Return_ratio(e)
head(e,15)
```



health and beauty aids products only occure in train, so we can ignore them in our model building
Concept stores, Other department products only get get returned in the test set
```{r}
dt.train <- data.table(train)
ScanCount_by_Visit_DepartmentDescription<-dt.train[,Quantity:=sum(ScanCount),by=list(VisitNumber,DepartmentDescription)]
ScanCount_by_Visit_DepartmentDescription = dcast(ScanCount_by_Visit_DepartmentDescription,VisitNumber~DepartmentDescription+Purchase_Ind)

dt.test <- data.table(test)
ScanCount_by_Visit_DepartmentDescription_test<-dt.test[,Quantity:=sum(ScanCount),by=list(VisitNumber,DepartmentDescription)]
ScanCount_by_Visit_DepartmentDescription_test = dcast(ScanCount_by_Visit_DepartmentDescription_test,VisitNumber~DepartmentDescription+Purchase_Ind)

a=names(ScanCount_by_Visit_DepartmentDescription) %in% names(ScanCount_by_Visit_DepartmentDescription_test)
b = names(ScanCount_by_Visit_DepartmentDescription)[!a]
c=names(ScanCount_by_Visit_DepartmentDescription_test) %in% names(ScanCount_by_Visit_DepartmentDescription)
d = names(ScanCount_by_Visit_DepartmentDescription_test)[!c]
b
d
```

let's look at records in Health and beauty aids records in train
we will replace heal and beauty aids to beauty
```{r}
train[which(train$VisitNumber %in% c(train[which(train$DepartmentDescription == "HEALTH AND BEAUTY AIDS"),]$VisitNumber)),]
```


Fineline number is a more defined product category made by Walmart.
```{r}
FinelineNumber_by_Deparment<-subset(dt.train,select = c('FinelineNumber','DepartmentDescription'))
FinelineNumber_by_Deparment<-FinelineNumber_by_Deparment[,Quantity:=length(unique(FinelineNumber)),by=list(DepartmentDescription),]
```


exaimne which transaction type exist in at
```{r}
xnam = c("ScanCount","Purchase_Ind","Weekday","TripType")
d<-aggregate(ScanCount ~ ., data=train[, xnam], FUN=sum,order('ScantCount'))
d = cast(d,DepartmentDescription~Purchase_Ind)
d = d[with(d, order(-Purchase)), ]
d[is.na(d)] <- 0
d = Calculate_Purchase_Return_ratio(d)
head(d,15)
```
